---
title: "Dennery lab scRNAseq"
author: "Joselynn"
date: "7/14/2020"
output: html_document
---


  
```{r Setup}
#install.packages('rstatix')
library(Seurat)
library(patchwork)
library(ggplot2)
library(dplyr)
library(future)
library(rsvd)
library(reshape2)
library(rstatix)
library(biomaRt)
library(stringr)
library(tidyr)
library(gridExtra)
library(data.table)
library(xlsx)
set.seed(61)
options(future.globals.maxSize = 4000 * 1024^2)
```

```{r Data import and merging}
# These lines don't work if run as part of a markdown code chunk, instead copy and paste them run into the rstudio console and them them there. them via the R console in RStudio (this seems to be an rstudio bug).
# Read in data
AirD7.data <- Read10X(data.dir ="AirD7out/filtered_feature_bc_matrix/")
AirD60.data <- Read10X(data.dir ="AirD60out/filtered_feature_bc_matrix/")
O2D7.data <- Read10X(data.dir ="O2D7out/filtered_feature_bc_matrix/")
O2D60.data <- Read10X(data.dir ="O2D60out/filtered_feature_bc_matrix/")
SD7.data <- Read10X(data.dir ="SD7out/filtered_feature_bc_matrix/")

# Create Seurat objects
#Include features (e.g. genes) detected in at least 3 cells and cells with at least 200 features.

AirD7.data <- CreateSeuratObject(counts = AirD7.data,  project = "AirD7",min.cells = 3, min.features = 200)
O2D7.data <- CreateSeuratObject(counts = O2D7.data,  project = "O2D7",min.cells = 3, min.features = 200)
AirD60.data <- CreateSeuratObject(counts = AirD60.data,  project = "AirD60",min.cells = 3, min.features = 200)
O2D60.data <- CreateSeuratObject(counts = O2D60.data,  project = "O2D60",min.cells = 3, min.features = 200)
SD7.data <- CreateSeuratObject(counts = SD7.data,  project = "SD7",min.cells = 3, min.features = 200)

# Merge the data
all_data <- merge(x = AirD7.data, y = c(O2D7.data, SD7.data, AirD60.data, O2D60.data), add.cell.ids = c("AirD7", "O2D7", "SD7", "AirD60", "O2D60"), project = 'dennery')

# Calculate the percentage of counts that belong to mitochondrial genes
all_data[["percent.mt"]] <- PercentageFeatureSet(all_data, pattern = "^mt-")

#Set levels
Idents(all_data) <- factor(Idents(all_data), levels = c("AirD7","O2D7","SD7", "AirD60", "O2D60"))
```

```{r QC Plots}
# QC Plots
VlnPlot(all_data, features = "nFeature_RNA")
VlnPlot(all_data, features = "nCount_RNA")
VlnPlot(all_data, features = "percent.mt")
#ggsave(file = '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/figures_and_tables/figures/qc_figures/nFeature_RNA_vlnplot.png', plot = VlnPlot(all_data, features = "nFeature_RNA"))
#ggsave(file = '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/figures_and_tables/figures/qc_figures/nCount_RNA_vlnplot.png', plot = VlnPlot(all_data, features = "nCount_RNA"))
#ggsave(file = '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/figures_and_tables/figures/qc_figures/percent.mt_vlnplot.png', plot = VlnPlot(all_data, features = "percent.mt"))

FeatureScatter(all_data, feature1 = "nCount_RNA", feature2 = "percent.mt")
FeatureScatter(all_data, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")

# one QC metric to look at is the number of unique genes detected per cell.
# low-quality/empty droplets will have very few genes or features.
# cell doublets or multiplets will have very high count.
# it looks like there's a cluster of cells in AirD7 and O2D7 where the features and counts are both low -- e.g., there's cells where there's only a few lowly-expressed genes.
# 8000 seems like a reasonable upper bound on the features.
#https://www.bioinformatics.babraham.ac.uk/training/10XRNASeq/seurat_workflow.html
qc.metrics <- dplyr::as_tibble(all_data[[c("nCount_RNA","nFeature_RNA","percent.mt", "orig.ident")]],rownames="Cell.Barcode") 
qc.metrics

mplot <- qc.metrics %>%
  arrange(percent.mt) %>%
  ggplot(aes(nCount_RNA,nFeature_RNA,colour=percent.mt)) + 
  geom_point() + 
  scale_color_gradientn(colors=c("black","blue","green2","red","yellow")) +
  ggtitle("QC metrics") +
  geom_hline(yintercept = 700) +
  geom_hline(yintercept = 8000) 

ident_plot <- qc.metrics %>%
  arrange(percent.mt) %>%
  ggplot(aes(nCount_RNA,nFeature_RNA,color=factor(orig.ident))) + 
  geom_point() + 
  scale_color_manual(values=c("#E69F00", "#56B4E9", "#009E73", "#F0E442")) +
  geom_hline(yintercept = 700) +
  geom_hline(yintercept = 8000) 

wrapped_filtering <- wrap_plots(mplot, ident_plot, ncol = 2, guides = "collect") + guides()

wrapped_filtering
```

```{r Filtering}

# Using a 5% mtDNA cutoff... as per https://www.biorxiv.org/content/10.1101/2020.02.20.958793v1.full.pdf
all_data_sub <- subset(all_data, subset = nFeature_RNA > 700 & nFeature_RNA < 8000 & percent.mt < 5)

all_data #20755 features across 16013 samples within 1 assay
all_data_sub #20755 features across 13199 samples within 1 assay 
```

```{r Post-filtering  QC plots}

FeatureScatter(all_data_sub, feature1 = "nCount_RNA", feature2 = "percent.mt")
FeatureScatter(all_data_sub, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
```


```{r Normalization}
# SCTransform to normalize, integrate, make some plots
#https://github.com/ChristophH/sctransform/issues/25 for explanation of why I'm not worried about 'iteration limit reached' warning messages.
all_data.list <- SplitObject(all_data_sub, split.by="orig.ident")

#sctransform
all_data.list <- lapply(all_data.list, function(x) {x <- SCTransform(x,verbose=FALSE)})
```

```{r Integration and clustering}
#all the integration steps
all_data.features <- SelectIntegrationFeatures(object.list=all_data.list,nfeatures=3000)
all_data.list <- PrepSCTIntegration(object.list=all_data.list, anchor.features=all_data.features, verbose=FALSE)
all_data.int.anchors <- FindIntegrationAnchors(all_data.list, normalization.method="SCT", anchor.features=all_data.features, verbose=FALSE)
all_data.combined <- IntegrateData(anchorset=all_data.int.anchors, normalization.method="SCT", verbose=FALSE)

DefaultAssay(all_data.combined) <- "integrated"

all_data.combined <- RunPCA(all_data.combined, features = VariableFeatures(object = all_data.combined))
all_data.combined <- FindNeighbors(all_data.combined, dims = 1:50)
all_data.combined <- FindClusters(all_data.combined, resolution = 2)
all_data.combined <- RunUMAP(all_data.combined, dims=1:50)
all_data.combined <- RunTSNE(all_data.combined, dims=1:50, nthreads=4, max_iter=2000, seed.use=35) 
#DimPlot(all_data.combined, reduction="tsne")
#all_data.combined <- RunTSNE(all_data.combined, dims=1:50, tsne.method = "FIt-SNE", nthreads=4, max_iter=2000, seed.use=35)  #Could not find the following programs: fast_tsne

```

```{r Import Mouse cell atlas data}
# MCA
#From https://satijalab.org/seurat/v3.1/mca.html vignette...

mca.matrix <- readRDS(file = "./MCA/MCA_merged_mat.rds")
mca.metadata <- read.csv("./MCA/MCA_All-batch-removed-assignments.csv")
mca.annotations <- read.csv("./MCA/MCA_CellAssignments.csv",row.names = 1)

# just interested in lung tissue
mca.annotations.lung <- mca.annotations %>% dplyr::filter(Tissue == 'Lung')
mca.metadata.lung <- mca.metadata %>% dplyr::filter(Tissue == 'Lung')

rownames(mca.annotations.lung) <- mca.annotations.lung$Cell.name
rownames(mca.metadata.lung) <- mca.metadata.lung$Cell.name

mca.metadata.annotations <- dplyr::inner_join(mca.metadata.lung, mca.annotations.lung, by = "Cell.name")
mca.metadata.annotations <- mca.metadata.annotations %>% dplyr::select(Cell.name, ClusterID.x, Tissue.x, Batch.x, Cell.Barcode.x, Annotation)
colnames(mca.metadata.annotations) <- c('Cell.name', 'ClusterID', 'Tissue', 'Batch', 'Cell.Barcode', 'Annotation')
rownames(mca.metadata.annotations) <- mca.metadata.annotations$Cell.name

mca.metadata.annotations <- mca.metadata.annotations %>% dplyr::select(-Cell.name)

mca.matrix.lung <- mca.matrix[,grepl("^Lung_", colnames(mca.matrix))]
mca <- CreateSeuratObject(counts = mca.matrix.lung, meta.data = mca.metadata.annotations, project = "MouseCellAtlas")

saveRDS(mca, file = "mca_sept_2021.rds")
saveRDS(all_data.combined, file = "all_data.combined_sept_2021.rds")
```

```{r Normalize mca, transfer anchors}
# Only keep annotated cells
mca <- SCTransform(mca)
# Leaves us with 242k cells

all_data.list$mca <- mca
atlas.features <- SelectIntegrationFeatures(all_data.list) #might need to tweak this 
atlas.anchors <- FindTransferAnchors(reference=mca, query=all_data.combined, normalization.method="SCT", npcs=82, dims=1:50, features=atlas.features, verbose=FALSE)

atlas.predictions <- TransferData(anchorset=atlas.anchors, refdata=mca$Annotation, dims=1:50)
all_data.combined <- AddMetaData(all_data.combined, metadata=atlas.predictions)
all_data.combined$atlas_predicted_clusters <- sapply(sapply(all_data.combined$predicted.id, function(x) strsplit(x, " - ")), `[`, 1)

all_data.combined@meta.data$atlas_predicted_clusters_short <- str_replace(all_data.combined@meta.data$atlas_predicted_clusters, pattern = "\\(Lung\\)", replacement = "")


saveRDS(mca, file = "mca_sept_2021.rds")
saveRDS(all_data.combined, file = "all_data.combined_sept_2021.rds")

```

```{r Fix 'Clara' to 'Club'}

all_data.combined$atlas_predicted_clusters <- str_replace(all_data.combined$atlas_predicted_clusters, pattern = 'Clara', replacement = 'Club')
all_data.combined$predicted.id <- str_replace(all_data.combined$predicted.id, pattern = 'Clara', replacement = 'Club')
all_data.combined$atlas_predicted_clusters_short <- str_replace(all_data.combined$atlas_predicted_clusters_short, pattern = 'Clara', replacement = 'Club')
colnames(all_data.combined@meta.data) <- str_replace(colnames(all_data.combined@meta.data), pattern = 'Clara', replacement = 'Club')

all_data.combined$orig.ident <- factor( x = all_data.combined$orig.ident, levels = c("AirD7", "O2D7", "SD7",  "AirD60", "O2D60"))

saveRDS(mca, file = "mca_sept_2021.rds")
saveRDS(all_data.combined, file = "all_data.combined_sept_2021.rds")


```

```{r Look at TSNE plots and compare mca vs seurat clusters}

p1 <- DimPlot(all_data.combined, reduction="tsne", group.by = "atlas_predicted_clusters") 
p1 <- LabelClusters(p1, id="atlas_predicted_clusters", repel=TRUE)
#ggsave(plot = p1, file = '2a1_nolabel.png', width = 10, height = 10, units = 'in')
p1

p2 <- DimPlot(all_data.combined, reduction="tsne", group.by = "seurat_clusters") #+ NoLegend()
p2 <- LabelClusters(p2, id="seurat_clusters", repel=TRUE)
#ggsave(plot = p2, file = '2a2_nolabel.png', width = 10, height = 10, units = 'in')
p2


#wrap_plots(list(p1,p2), ncol = 2)

p3 <- DimPlot(all_data.combined, reduction="tsne", group.by = "atlas_predicted_clusters_short", split.by = "orig.ident") 
p3 <- LabelClusters(p3, id="atlas_predicted_clusters_short", repel=TRUE)
p3

p4 <- DimPlot(all_data.combined, reduction="tsne", group.by = "seurat_clusters", split.by = "orig.ident") 
#p4 <- LabelClusters(p4, id="seurat_clusters", repel=TRUE)
p4
```

```{r Get more info per each gene}

# get a list of all the genes in the Seurat object.
all_genes <- data.frame(rownames(all_data.combined))
all_genes$mgi_symbol <- all_genes$rownames.all_data.combined.
all_genes$rownames.all_data.combined. <- NULL

mouse_biomart <- useMart("ensembl", dataset = "mmusculus_gene_ensembl")
mouse_biomart_data <- getBM(mart = mouse_biomart, attributes = c("mgi_description", "mgi_symbol"))
all_genes_bm <- right_join(mouse_biomart_data, all_genes)

saveRDS(all_genes_bm, "all_genes_bm.rds")
#all_genes_bm <- readRDS("/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/all_genes_bm.RDS")

```

```{r Try making some more compelling featureplots for the genes Hongwei suggested}

all_genes_bm <- readRDS('all_genes_bm.RDS')

#Please see below the gene list which can be used for identification of senescence. 
#Gene list of senescence:
#Cdkn1a (encoding p21 protein), cdkn2a (encoding p14, p16, and p19 proteins), cdkn2b (encoding p15 protein), p53, Rb, e2f.
#We may also want to see genes (ki67, pcna) of proliferation, which is usually decreased in senescent cells.
#Some genes including sirt1, sirt3, sirt6, lmnb1, egr1, ezh2 are also associated with senescence
#Genes of senescence-associated secretory phenotype (SASP) include IL-1, IL-6, IL-10, PAI1, TNF-α, CXCL1/KC, CXCL2, CCL3, CCL2, and CCL2/MCP-1, MMP3, 
cdkn <- grep('cdkn', all_genes_bm$mgi_symbol, ignore.case = T, value = TRUE)
p53 <- grep('p53', all_genes_bm$mgi_symbol, ignore.case = T, value = TRUE)
#didn't find 'Rb'
e2f <- grep('e2f', all_genes_bm$mgi_symbol, ignore.case = T, value = TRUE)
ki67 <- grep('ki67', all_genes_bm$mgi_symbol, ignore.case = T, value = TRUE)
pcna <- grep('pcna', all_genes_bm$mgi_symbol, ignore.case = T, value = TRUE)
#didn't find 'Sirt'
lmnb1 <- grep('lmnb1', all_genes_bm$mgi_symbol, ignore.case = T, value = TRUE)
egr1 <- grep('egr1', all_genes_bm$mgi_symbol, ignore.case = T, value = TRUE)
ezh2 <- grep('ezh2', all_genes_bm$mgi_symbol, ignore.case = T, value = TRUE)
il6 <-'Il6'
il1 <- c('Il1a','Il1b')  
il10 <- c('Il10rb', 'Il10ra')
#serpine is synonym for pai1 http://www.informatics.jax.org/marker/MGI:97608
serpine1 <- grep('Serpine1', all_genes_bm$mgi_symbol, ignore.case = T, value = TRUE)
tnfs <- grep('tnf', all_genes_bm$mgi_symbol, ignore.case = T, value = TRUE)
cxc <- grep('cxc', all_genes_bm$mgi_symbol, ignore.case = T, value = TRUE)
ccl <- grep('ccl', all_genes_bm$mgi_symbol, ignore.case = T, value = TRUE)
mmp <- grep('mmp', all_genes_bm$mgi_symbol, ignore.case = T, value = TRUE)

DefaultAssay(all_data.combined) <- 'SCT'
Idents(all_data.combined) <- 'seurat_clusters'
                        
O2D7_seurat <- subset(all_data.combined, orig.ident == 'O2D7')
AirD7_seurat <- subset(all_data.combined, orig.ident == 'AirD7')
SD7_seurat <- subset(all_data.combined, orig.ident == 'SD7')
O2D60_seurat <- subset(all_data.combined, orig.ident == 'O2D60')
AirD60_seurat <- subset(all_data.combined, orig.ident == 'AirD60')


Idents(O2D7_seurat) <- 'seurat_clusters'
Idents(AirD7_seurat) <- 'seurat_clusters'
Idents(SD7_seurat ) <- 'seurat_clusters'
Idents(O2D60_seurat) <- 'seurat_clusters'
Idents(AirD60_seurat) <- 'seurat_clusters'


featureplot_grid <- function(feats, pwidth, pheight, plotfile){
  p_aird7 <- FeaturePlot(AirD7_seurat, 
                            label = TRUE, 
                            label.size = 6, 
                            repel = F, 
                            sort = TRUE, 
                            features = feats, 
                            ncol = length(feats), 
                            reduction = 'tsne') &
  scale_colour_gradientn(limits= c(0,8), 
                         colors = c('bisque', 'darkred')) &
  labs(subtitle = 'Air D7')

p_o2d7 <- FeaturePlot(O2D7_seurat, 
                            label = TRUE, 
                            label.size = 6, 
                            repel = F, 
                            sort = TRUE, 
                            features = feats, 
                            ncol = length(feats), 
                            reduction = 'tsne') &
  scale_colour_gradientn(limits= c(0,8), 
                         colors = c('bisque', 'darkred')) &
  labs(subtitle = 'O2 D7')

p_sd7 <- FeaturePlot(SD7_seurat, 
                            label = TRUE, 
                            label.size = 6, 
                            repel = F, 
                            sort = TRUE, 
                            features = feats, 
                            ncol = length(feats), 
                            reduction = 'tsne') &
  scale_colour_gradientn(limits= c(0,8), 
                         colors = c('bisque', 'darkred')) &
  labs(subtitle = 'S D7')

p_aird60 <- FeaturePlot(AirD60_seurat, 
                            label = TRUE, 
                            label.size = 6, 
                            repel = F, 
                            sort = TRUE, 
                            features = feats, 
                            ncol = length(feats), 
                            reduction = 'tsne') &
  scale_colour_gradientn(limits= c(0,8), 
                         colors = c('bisque', 'darkred')) &
  labs(subtitle = 'Air D60')

p_o2d60 <- FeaturePlot(O2D60_seurat, 
                            label = TRUE, 
                            label.size = 6, 
                            repel = F, 
                            sort = TRUE, 
                            features = feats, 
                            ncol = length(feats), 
                            reduction = 'tsne') &
  scale_colour_gradientn(limits= c(0,8), 
                         colors = c('bisque', 'darkred')) &
  labs(subtitle = 'O2 D60')


plots <- p_aird7 / p_o2d7 / p_sd7 / p_aird60 / p_o2d60

ggsave(plotfile, device = 'png', width = pwidth, height = pheight)
  
return(plots)
}

featureplot_grid(cdkn, pwidth = 30, pheight = 30, 'p_cdkn_fp.png')
featureplot_grid(p53, pwidth = 6, pheight = 30, 'p_p53_fp.png')
featureplot_grid(e2f, pwidth = 15, pheight = 30, 'p_e2f_fp.png')
featureplot_grid(ki67, pwidth = 6, pheight = 30, 'p_ki67_fp.png')
featureplot_grid(pcna, pwidth = 6, pheight = 30, 'p_pcna_fp.png')
featureplot_grid(lmnb1, pwidth = 6, pheight = 30, 'p_lmnb1_fp.png')
featureplot_grid(egr1, pwidth = 6, pheight = 30, 'p_egr1_fp.png')
featureplot_grid(ezh2, pwidth = 6, pheight = 30, 'p_ezh2_fp.png')
featureplot_grid(il6, pwidth = 6, pheight = 30, 'p_il6_fp.png')
featureplot_grid(il1, pwidth = 15, pheight = 30, 'p_il1_fp.png')
featureplot_grid(il10, pwidth = 15, pheight = 30, 'p_il10_fp.png')
featureplot_grid(serpine1, pwidth = 6, pheight = 30, 'p_serpine1_fp.png')
featureplot_grid(tnfs, pwidth = 40, pheight = 30, 'p_tnfs_fp.png')
featureplot_grid(cxc, pwidth = 40, pheight = 30, 'p_cxc_fp.png')
featureplot_grid(mmp, pwidth = 40, pheight = 30, 'p_mmp_fp.png')
```

```{r Make a dotplot of Hongwei's genes of interest}

Idents(all_data.combined) <- 'orig.ident'

features_list <- c(cdkn, p53, e2f, ki67, pcna, lmnb1, egr1, ezh2, il6, il1, il10, serpine1, tnfs, cxc, mmp)

all_dotplot <-  DotPlot(all_data.combined, assay = 'SCT',features = features_list ) & theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) & coord_flip()

ggsave('all_dotplot.png', plot = all_dotplot, width = 6)

p_cdkn_dot <- DotPlot(all_data.combined, assay = 'SCT',features = cdkn)
p_p53_dot <- DotPlot(all_data.combined, assay = 'SCT',features = p53)
p_e2f_dot <- DotPlot(all_data.combined, assay = 'SCT',features = e2f)
p_ki67_dot <- DotPlot(all_data.combined, assay = 'SCT',features = ki67)
p_pcna_dot <- DotPlot(all_data.combined, assay = 'SCT',features = pcna)
p_lmnb1_dot <- DotPlot(all_data.combined, assay = 'SCT',features = lmnb1)
p_egr1_dot <- DotPlot(all_data.combined, assay = 'SCT',features = egr1)
p_ezh2_dot <- DotPlot(all_data.combined, assay = 'SCT',features = ezh2)
p_il6_dot <- DotPlot(all_data.combined, assay = 'SCT',features = il6)
p_il1_dot <- DotPlot(all_data.combined, assay = 'SCT',features = il1)
p_il10_dot <- DotPlot(all_data.combined, assay = 'SCT',features = il10)
p_serpine1_dot <- DotPlot(all_data.combined, assay = 'SCT',features = serpine1)

p_tnfs_dot <- DotPlot(all_data.combined, assay = 'SCT',features = tnfs)
p_cxc_dot <- DotPlot(all_data.combined, assay = 'SCT',features = cxc)
p_mmp_dot <- DotPlot(all_data.combined, assay = 'SCT',features = mmp)

ggsave(filename = 'p_cdkn_dot.png', plot = p_cdkn_dot, device = 'png', width = 12)
ggsave(filename = 'p_p53_dot.png', plot = p_p53_dot, device = 'png', width = 12)
ggsave(filename = 'p_e2f_dot.png', plot = p_e2f_dot, device = 'png', width = 12)
ggsave(filename = 'p_ki67_dot.png', plot = p_ki67_dot, device = 'png', width = 12)
ggsave(filename = 'p_pcna_dot.png', plot = p_pcna_dot, device = 'png', width = 12)
ggsave(filename = 'p_lmnb1_dot.png', plot = p_lmnb1_dot, device = 'png', width = 12)
ggsave(filename = 'p_egr1_dot.png', plot = p_egr1_dot, device = 'png', width = 12)
ggsave(filename = 'p_ezh2_dot.png', plot = p_ezh2_dot, device = 'png', width = 12)
ggsave(filename = 'p_il6_dot.png', plot = p_il6_dot, device = 'png', width = 12)
ggsave(filename = 'p_il1_dot.png', plot = p_il1_dot, device = 'png', width = 12)
ggsave(filename = 'p_il10_dot.png', plot = p_il10_dot, device = 'png', width = 12)
ggsave(filename = 'p_serpine1_dot.png', plot = p_serpine1_dot, device = 'png', width = 12)
ggsave(filename = 'p_tnfs_dot.png', plot = p_tnfs_dot, device = 'png', width = 12)
ggsave(filename = 'p_cxc_dot.png', plot = p_cxc_dot, device = 'png', width = 12)
ggsave(filename = 'p_mmp_dot.png', plot = p_mmp_dot, device = 'png', width = 12)

#SCT counts slot is the corrected counts
Idents(SD7_seurat) <- SD7_seurat$seurat_clusters

foi_norm <- AverageExpression(SD7_seurat, assays = 'SCT', features = c(features_list), slot = 'counts')
write.table(foi_norm, file = 'sd7_normalized_filtered_counts.csv')

sd7_norm <- AverageExpression(SD7_seurat, assays = 'SCT', slot = 'counts')

write.table(sd7_norm, file = 'sd7_normalized_counts.csv')


```

```{r compare clusters and mca annotations dim reduction}
atlas_plot <- DimPlot(all_data.combined, reduction="tsne", group.by = "atlas_predicted_clusters_short", split.by = 'orig.ident', label = TRUE, label.size = 3, repel = T) 
seurat_plot <-  DimPlot(all_data.combined, reduction="tsne", group.by = "seurat_clusters", split.by = 'orig.ident', label = TRUE, label.size = 3, repel = T) 

seurat_v_atlas <- seurat_plot / atlas_plot

ggsave(filename = 'seurat_v_atlas.png', plot = seurat_v_atlas, device = 'png', width = 30)
```
 
```{r This section is where we find markers distinguishing the clusters only in SD7}


Idents(SD7_seurat) <- SD7_seurat$seurat_clusters
# find markers for every cluster compared to all remaining cells, report only the positive ones
SD7_seurat.markers <- FindAllMarkers(SD7_seurat, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, assay = 'SCT')

#saveRDS(SD7_seurat.markers, file = "SD7_seurat.markers.rds")
#saveRDS(SD7_seurat, file = "SD7_seurat.rds")

# Add the gene information to each cluster..
SD7_seurat.markers <- left_join(SD7_seurat.markers, all_genes_bm, by = c('gene' = 'mgi_symbol'))

#get distinct markers
SD7_seurat.markers_distinct <- distinct(SD7_seurat.markers, gene, .keep_all = T)

#all_data.combined <- readRDS("all_data.combined_sept_2021.rds")
#SD7_seurat.markers <- readRDS("SD7_seurat.markers.rds")
#SD7_seurat <- readRDS("SD7_seurat.rds")

#split up the markers by cluster..
cluster_levels <- unique(SD7_seurat.markers_distinct$cluster) 
SD7_seurat.markers.split <- lapply(seq_along(cluster_levels), function(i) SD7_seurat.markers_distinct %>% dplyr::filter(cluster == cluster_levels[i]))


#install.packages("writexl")
library(writexl)

sheet_names <- c(paste0('cluster', '_', cluster_levels))

names(SD7_seurat.markers.split) <- paste0('seurat_', unique(top5_markers$cluster))

write_xlsx(SD7_seurat.markers.split, "markers_per_cluster_SD7.xlsx")

#doing some filtering to look at just the 'top markers', but probably want to only use these for making figures.
top5_markers <- SD7_seurat.markers_distinct  %>% group_by(cluster) %>% top_n(n = 5, wt = avg_logFC)

top5_markers.split <- lapply(seq_along(unique(top5_markers$cluster)), function(i) top5_markers %>% dplyr::filter(cluster == unique(top5_markers$cluster)[i]))
  
marker_dotplots <- lapply(seq_along(top5_markers.split), function(i) DotPlot(SD7_seurat, features = c((top5_markers.split[[i]])$gene)) & labs(title = paste0('seurat_', (top5_markers.split[[i]])$cluster)))

```
```{r}
(marker_dotplots[[1]] + marker_dotplots[[2]] + marker_dotplots[[3]]) / (marker_dotplots[[4]] + marker_dotplots[[5]] + marker_dotplots[[6]]) 
```


```{r}
ggsave(filename = 'marker_dotplots.png', width = 30, height = 30, device = 'png', plot = (marker_dotplots[[1]] + marker_dotplots[[2]] + marker_dotplots[[3]] + marker_dotplots[[4]] + marker_dotplots[[5]] + marker_dotplots[[6]] + marker_dotplots[[7]] + marker_dotplots[[8]] + marker_dotplots[[9]] + marker_dotplots[[10]] + 
marker_dotplots[[11]] + marker_dotplots[[12]] + marker_dotplots[[13]] + marker_dotplots[[14]] + marker_dotplots[[15]] + marker_dotplots[[16]] + marker_dotplots[[17]] + marker_dotplots[[18]] + marker_dotplots[[19]] + marker_dotplots[[20]] + marker_dotplots[[21]] + marker_dotplots[[22]]))
```



```{r}

DefaultAssay(SD7_seurat) <- 'SCT'

marker_dotplots <-  DotPlot(SD7_seurat, features = top5_markers$gene, group.by = 'seurat_clusters') & 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) & 
  geom_vline(xintercept = 5.55) &
  geom_vline(xintercept = 10.55) &
  geom_vline(xintercept = 15.55) &
  geom_vline(xintercept = 20.55) &
  geom_vline(xintercept = 25.55) &
  geom_vline(xintercept = 30.55) &
  geom_vline(xintercept = 35.55) &
  geom_vline(xintercept = 40.55) &
  geom_vline(xintercept = 45.55) &
  geom_vline(xintercept = 50.55) &
  geom_vline(xintercept = 55.55) &
  geom_vline(xintercept = 60.55) &
  geom_vline(xintercept = 65.55) &
  geom_vline(xintercept = 70.55) &
  geom_vline(xintercept = 75.55) &
  geom_vline(xintercept = 80.55) &
  geom_vline(xintercept = 85.55) &
  geom_vline(xintercept = 90.55) &
  geom_vline(xintercept = 95.55) &
  geom_vline(xintercept = 100.55) &
  geom_vline(xintercept = 105.55) &
  geom_vline(xintercept = 110.55) &
  geom_hline(yintercept = 1.5) &
  geom_hline(yintercept = 2.5) &
  geom_hline(yintercept = 3.5) &
  geom_hline(yintercept = 4.5) &
  geom_hline(yintercept = 5.5) &
  geom_hline(yintercept = 6.5) &
  geom_hline(yintercept = 7.5) &
  geom_hline(yintercept = 8.5) &
  geom_hline(yintercept = 9.5) &
  geom_hline(yintercept = 10.5) &
  geom_hline(yintercept = 11.5) &
  geom_hline(yintercept = 12.5) &
  geom_hline(yintercept = 13.5) &
  geom_hline(yintercept = 14.5) &
  geom_hline(yintercept = 15.5) &
  geom_hline(yintercept = 16.5) &
  geom_hline(yintercept = 17.5) &
  geom_hline(yintercept = 18.5) &
  geom_hline(yintercept = 19.5) &
  geom_hline(yintercept = 20.5) &
  geom_hline(yintercept = 21.5) &
  geom_hline(yintercept = 22.5) &
  geom_hline(yintercept = 23.5) &
  geom_hline(yintercept = 24.5) &
  geom_hline(yintercept = 25.5) &
  geom_hline(yintercept = 26.5) &
  geom_hline(yintercept = 27.5) &
  geom_hline(yintercept = 28.5) &
  geom_hline(yintercept = 29.5) &
  geom_hline(yintercept = 30.5) &
  geom_hline(yintercept = 31.5) &
  geom_hline(yintercept = 32.5) &
  geom_hline(yintercept = 33.5)




#Results are saved in a new assay (named SCT by default) with counts being (corrected) counts, data being log1p(counts), scale.data being pearson residuals; sctransform::vst intermediate results are saved in misc slot of new assay. 

FeaturePlot(SD7_seurat, 
            label = TRUE, 
            label.size = 6, 
            repel = F, 
            sort = TRUE, 
            features = (top5_markers %>% dplyr::filter(cluster == 0))$gene, 
            ncol = 5, 
            reduction = 'tsne',
            slot = 'counts') &
  scale_colour_gradientn(limits= c(0,8), colors = c('bisque', 'darkred')) &
  labs(subtitle = 'SD7')


DotPlot(SD7_seurat, 
        assay = 'SCT',
        features = (top5_markers %>% dplyr::filter(cluster == 8))$gene) & 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) & 
  coord_flip()
```

```{r Table comparing the numbers of cells in each combination of seurat cluster and cell type annotation}
cell_types_per_cluster_df  <- data.frame(atlas = SD7_seurat$atlas_predicted_clusters_short, seurat = SD7_seurat$seurat_clusters)
cell_types_per_cluster <- as.data.frame(table(cell_types_per_cluster_df$atlas, cell_types_per_cluster_df$seurat))
cell_types_per_cluster$atlas <- cell_types_per_cluster$Var1
cell_types_per_cluster$seurat <- cell_types_per_cluster$Var2
cell_types_per_cluster$Var1 <- NULL
cell_types_per_cluster$Var2 <- NULL
cell_types_per_cluster <- cell_types_per_cluster %>% arrange(desc(Freq))
cell_types_per_cluster <- cell_types_per_cluster %>% filter(Freq > 0)
cell_types_per_cluster <- cell_types_per_cluster %>% arrange(desc(atlas))
#write.csv(cell_types_per_cluster, 'cell_types_per_cluster.csv', row.names= F)

cell_types_per_cluster$number_of_cells <- cell_types_per_cluster$Freq
cell_types_per_cluster$Freq <- NULL
cell_types_per_cluster_plot <- ggplot(cell_types_per_cluster, aes(x = atlas, y = seurat, fill = number_of_cells)) +
  coord_equal()+
  geom_tile(colour = 'black') + 
  xlab('Cell Annotation') +
  ylab('Seurat Cluster') +
  theme(panel.grid = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks = element_blank()) +
  scale_fill_gradientn(
    colors = c("white", "darkred"),
    limits = range(cell_types_per_cluster$number_of_cells)) +
  theme(panel.grid = element_blank(),
        axis.text.x=element_text(angle=90, hjust = 1, vjust = .4),
        axis.ticks = element_blank(),
        text = element_text(size=20)) 
```
```{r Table of how many cells in each cell type category}
cell_types_freq_df <- data.frame(ident = SD7_seurat$orig.ident, atlas = SD7_seurat$atlas_predicted_clusters_short)
cell_types_freq <- as.data.frame(table(cell_types_freq_df$ident, cell_types_freq_df$atlas))
cell_types_freq$atlas <- cell_types_freq$Var2
cell_types_freq$Var2 <- NULL
cell_types_freq$ident <- cell_types_freq$Var1
cell_types_freq$Var1 <- NULL

cell_types_freq <- cell_types_freq %>% arrange(desc(Freq))
cell_types_freq <- cell_types_freq %>% filter(Freq > 0)
#write.csv(cell_types_freq, 'cell_types_freq.csv', row.names= F)


cell_types_freq$number_of_cells <- cell_types_freq$Freq
cell_types_freq$Freq <- NULL

cell_types_freq_plot <- ggplot(cell_types_freq, aes(x = ident, y = atlas, fill = number_of_cells)) +
  coord_equal()+
  geom_tile(colour = 'black') + 
  xlab('Ident') +
  ylab('Cell state') +
  theme(panel.grid = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks = element_blank()) +
  scale_fill_gradientn(
    colors = c("white", "red"),
    limits = range(cell_types_freq$number_of_cells)) +
  geom_text(aes(label = number_of_cells)) +
  theme(panel.grid = element_blank(),
        axis.text.x=element_text(angle=90, hjust = 1, vjust = .4),
        axis.ticks = element_blank(),
        text = element_text(size=20))

```
```{r Table of how many cells are in each seurat cluster}
clusters_freq_df <- data.frame(ident = SD7_seurat$orig.ident, seurat = SD7_seurat$seurat_clusters)
clusters_freq <- as.data.frame(table(clusters_freq_df$ident, clusters_freq_df$seurat))
clusters_freq$seurat <- clusters_freq$Var2
clusters_freq$Var1 <- NULL
clusters_freq$Var2 <- NULL
clusters_freq <- clusters_freq %>% arrange(desc(Freq))
clusters_freq$Number_of_cells <- clusters_freq$Freq
#write.csv(clusters_freq, 'clusters_freq.csv', row.names= F)
```


```{r}




ggsave('cell_types_per_cluster_plot.png', device = 'png')
#ggsave(plot = conc_plot2, file = 'figure2b1.png', width = 10, height = 15, units = 'in')

```













```


# AT1, AT2, and B cells are in similar clusters (does that make sense? are they similar cell types?)
# AT1 cells are in cluster 25, top markers: Ndst1, Mal2, Tead1, Agrn, Dag1, between 35 and 52 cells in each experiment
# AT2 cells are in a few clusters -- 3, 11, 13, 22, and 27. 22 and 27 have other cell types too (B cells, AT1 cells, alveolar macrophages), but are mainly AT2.
# Cluster 3 markers: Cat, Irx3, Pla2g1b, Ctsh, Elf5
# Cluster 11 markers: Kcnc3, LOC115487417, Ppp1r9a, Gm33122, St6galnac2
# Cluster 13 markers: Fasn, Tgoln1, Il33, Lrg2, Lrp2
# Cluster 22 markers: 



```{r How many cells are in each cell type?}
DefaultAssay(all_data.combined) <- 'SCT'

all_data.combined_meta.data <- tibble(all_data.combined@meta.data) 

counts_per_cluster_and_ident <- all_data.combined_meta.data %>% group_by(seurat_clusters, orig.ident) %>% tally() 
counts_per_cluster_and_ident_spread <- counts_per_cluster_and_ident %>% tidyr::spread(key=orig.ident, value=n)
nrow(all_data.combined_meta.data)
colSums(Filter(is.numeric, counts_per_cluster_and_ident_spread), na.rm = T)


counts_per_annot_and_ident <- all_data.combined_meta.data %>% group_by(predicted.id, orig.ident) %>% tally() 
counts_per_annot_and_ident_spread <- counts_per_annot_and_ident %>% tidyr::spread(key=orig.ident, value=n)
nrow(all_data.combined_meta.data)
colSums(Filter(is.numeric, counts_per_annot_and_ident_spread), na.rm = T)

table(all_data.combined@meta.data$seurat_clusters, all_data.combined@meta.data$orig.ident)
table(all_data.combined@active.ident, all_data.combined@meta.data$orig.ident)
table(counts_per_annot_and_ident_spread)

#https://github.com/satijalab/seurat/issues/738

count_plot_seurat <- ggplot(counts_per_cluster_and_ident, aes(x = orig.ident, y = seurat_clusters, fill = n)) + 
  geom_tile() +
  theme(panel.grid = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks = element_blank()) +
  geom_text(aes(label = n)) +
  xlab("") +
  scale_fill_gradientn(
    colors = c("white", "steelblue3"),
    limits = range(counts_per_cluster_and_ident$n)) +
  theme_minimal()

count_plot_mca <- ggplot(counts_per_annot_and_ident, aes(x = orig.ident, y = predicted.id, fill = n)) + 
  geom_tile() +
  theme(panel.grid = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks = element_blank()) +
  geom_text(aes(label = n)) +
  xlab("") +
  scale_fill_gradientn(
    colors = c("white", "steelblue3"),
    limits = range(counts_per_annot_and_ident$n)) +
  theme_minimal()

#write.table((table(all_data.combined@meta.data$seurat_clusters, all_data.combined@meta.data$orig.ident)), "/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/cells_per_seurat_cluster.txt", sep = '\t')

#write.table((table(all_data.combined@meta.data$predicted.id, all_data.combined@meta.data$orig.ident)), "/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/cells_per_mca_cluster_annot.txt", sep = '\t')
```


```{r}
#saveRDS(all_data.combined, file = "/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/all_data.combined.RDS")
#all_data.combined <- readRDS(file = "/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/all_data.combined.RDS")
```

```{r Differential Expression}
# It seems like folks are interested in a few clusters:
# Cluster 25 - AT1 cells
# Clusters 3, 11, 13, 22, and 27 are the AT2 cell types (22 and 27 have other cell types too).
# Ciliated cells are in Cluster 28
# Club cells are in Cluster 39

levels(all_data.combined$orig.ident)

#make a new column in all_data.combined@meta.data that combines the seurat cluster and orig.ident
all_data.combined@meta.data$ident_clust <- paste0(all_data.combined@meta.data$orig.ident, "-", all_data.combined@meta.data$seurat_clusters)

all_data.combined@meta.data$treat <- tidyr::separate(data = all_data.combined@meta.data, 
                col = orig.ident,
                into = c('treat','time'),
                sep = 'D',
                remove = F)$treat


all_data.combined@meta.data$time <- tidyr::separate(data = all_data.combined@meta.data, 
                col = orig.ident,
                into = c('treat','time'),
                sep = 'D',
                remove = F)$time
```

```{r Look at marker genes across clusters e.g., which genes are most important for distinguishing the cell clusters -- we were using the integrated data previously, but I think using the 'SCT' normalized data is a better idea.}
Idents(all_data.combined) <- all_data.combined$seurat_clusters
DefaultAssay(all_data.combined) <- 'SCT'

# find markers for every cluster compared to all remaining cells, report only the positive ones
all_data.combined.markers.sct <- FindAllMarkers(all_data.combined, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, assay = 'SCT')

# Add the gene information to each cluster..
all_data.combined.markers.sct <- left_join(all_data.combined.markers.sct, all_genes_bm, by = c('gene' = 'mgi_symbol'))

#split up the markers by cluster..
cluster_levels <- levels(all_data.combined.markers.sct$cluster)
all_data.combined.markers.sct.split <- lapply(seq_along(cluster_levels), function(i) all_data.combined.markers.sct %>% dplyr::filter(cluster == cluster_levels[i]))

#per_cluster_markers <- all_data.combined.markers.sct.split

#saveRDS(per_cluster_markers, '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/per_cluster_markers.Rds')

per_cluster_markers <- readRDS('/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/per_cluster_markers.Rds')
```


```{r}
Idents(all_data.combined) <- all_data.combined$orig.ident
DefaultAssay(all_data.combined) <- 'SCT'

# This chunk will take the Seurat object as an input and filter is such that each seurat cluster is in its own seurat object and then findallmarkers, then make dotplots of the top markers across the experimental conditions

#set the levels
clusters <- levels(all_data.combined@meta.data$seurat_clusters)

construct <- unlist(lapply(seq_along(clusters), function(i) paste0("all_data.combined_", clusters[i], "<- subset(all_data.combined, subset = seurat_clusters =='", clusters[i], "')")))

seurat.list <- lapply(seq_along(construct), function(i) eval(parse(text = construct[i])))

names(seurat.list) <- c(levels(all_data.combined@meta.data$seurat_clusters))

seurat.list.markers <- lapply(seurat.list, FindAllMarkers, min.pct = 0.25, logfc.threshold = 0.25, assay = 'SCT')

#Add the gene explanations to each file..
seurat.list.markers.ids <- lapply(seq_along(seurat.list.markers), function(i) left_join(seurat.list.markers[[i]], all_genes_bm, by = c('gene' = 'mgi_symbol')))

per_condition_markers <- seurat.list.markers.ids
#saveRDS(per_condition_markers, '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/per_condition_markers.Rds')

names(per_condition_markers) <- names(seurat.list)

#write the findallmarkers outputs for each cluster to a file
#lapply(seq_along(seurat.list.markers.ids), function(i) write.csv(seurat.list.markers.ids[[i]], file = paste0("/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/figures_and_tables/tables/condition_markers/cluster_", i-1, "_per_condition_markers.csv"), row.names = F))

```
```{r}
DefaultAssay(all_data.combined) <- 'SCT'
FeaturePlot(all_data.combined, features = 'Malat1', split.by = 'orig.ident', reduction = 'tsne', slot = 'data', cols = )

```



```{r Re-import the per-condition markers and try to collate them into one table that focuses on the O2D60 data}
#first, add a column to each object in the list, the column should just be its own index or name
per_condition_markers_bind <- rbindlist(per_condition_markers, idcol = "seurat_cluster")

per_condition_markers_bind_sort <- per_condition_markers_bind %>% dplyr::arrange(factor(cluster, levels = c('O2D60','AirD60', 'O2D7', 'AirD7')))

per_condition_markers_bind_sort <- dplyr::left_join(per_condition_markers_bind_sort, all_genes_bm, by = c('gene' = 'mgi_symbol'))

write.csv(per_condition_markers_bind_sort, file = '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/per_condition_markers_bind_sort.csv', row.names = F)

write.xlsx(per_condition_markers_bind_sort, file = '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/per_condition_markers_bind_sort.xlsx', row.names = F)
```



```{r Looking at overall condition markers (rather than per-cluster), looking at featureplots..}

Idents(all_data.combined) <- all_data.combined$orig.ident
DefaultAssay(all_data.combined) <- 'SCT'

all_condition_markers <- FindAllMarkers(all_data.combined, min.pct = 0.25, logfc.threshold = 0.25, assay = 'SCT')

all_condition_markers_sort <- all_condition_markers %>% dplyr::arrange(factor(cluster, levels = c('O2D60','AirD60', 'O2D7', 'AirD7')))
all_condition_markers_sort %>% dplyr::filter(cluster == 'O2D60')
-----
  
Idents(all_data.combined) <- all_data.combined$seurat_clusters
  
features <- c("Cftr", "Spock2", "Malat1", "Tkt", "Alas1")
DotPlot(all_data.combined, features = features, group.by = 'seurat_clusters', assay = 'SCT') + RotatedAxis()
    
rownames(all_data.combined@assays$SCT@counts)

all_data.combined$seurat_clusters
```


```{r}

get_top_markers <- function(seurat_markers){
  output <- seurat_markers %>% group_by(cluster)  %>% top_n(n = 10, wt = avg_logFC)
  return(output)
}

seurat.list.top.markers <- lapply(seurat.list.markers, get_top_markers)

#seurat_list_dotplots <- lapply(seq_along(seurat.list), function(i) DotPlot(seurat.list[[i]], features = c(unique(seurat.list.top.markers[[i]]$gene))) + RotatedAxis() + ggtitle(paste0("Seurat Cluster ", names(seurat.list[i]))))

#lapply(seq_along(seurat_list_dotplots), function(i) ggsave(plot =) seurat_list_dotplots[[i]], file = paste0("/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/figures/cluster_", i-1, "_dotplot.pdf"), width = 11,  height = 7))
```

```{r}
#3, 11, 13, 22, 27

cluster_3_markers <- per_cluster_markers[[4]]
cluster_11_markers <- per_cluster_markers[[12]]
cluster_13_markers <- per_cluster_markers[[14]]
cluster_22_markers <- per_cluster_markers[[23]]
cluster_27_markers <- per_cluster_markers[[28]]
```

```{r}
#make some lists..
cluster_3_compare <- cluster_11_markers %>%
  full_join(cluster_13_markers, by = 'gene') %>%
  full_join(cluster_22_markers, by = 'gene') %>%
  full_join(cluster_27_markers, by = 'gene')

cluster_11_compare <- cluster_3_markers %>%
  full_join(cluster_13_markers, by = 'gene') %>%
  full_join(cluster_22_markers, by = 'gene') %>%
  full_join(cluster_27_markers, by = 'gene')

cluster_13_compare <- cluster_11_markers %>%
  full_join(cluster_3_markers, by = 'gene') %>%
  full_join(cluster_22_markers, by = 'gene') %>%
  full_join(cluster_27_markers, by = 'gene')

cluster_22_compare <- cluster_11_markers %>%
  full_join(cluster_3_markers, by = 'gene') %>%
  full_join(cluster_13_markers, by = 'gene') %>%
  full_join(cluster_27_markers, by = 'gene')

cluster_27_compare <- cluster_11_markers %>%
  full_join(cluster_3_markers, by = 'gene') %>%
  full_join(cluster_13_markers, by = 'gene') %>%
  full_join(cluster_22_markers, by = 'gene')

#markers in cluster 3 but not in 11, 13, 22, or 27
cluster_3_specific <- anti_join(cluster_3_markers, cluster_3_compare, by="gene") 
cluster_11_specific <- anti_join(cluster_11_markers, cluster_11_compare, by="gene") 
cluster_13_specific <- anti_join(cluster_13_markers, cluster_13_compare, by="gene") 
cluster_22_specific <- anti_join(cluster_22_markers, cluster_22_compare, by="gene") 
cluster_27_specific <- anti_join(cluster_27_markers, cluster_27_compare, by="gene") 

```

```{r}
#Compare to supplemental data
supp_table<-read.csv('/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/41586_2018_BFnature25786_MOESM2_ESM.csv')
supp_table <- left_join(supp_table, all_genes_bm, by = c('SYMBOL' = 'mgi_symbol'))
```

```{r}
cluster_3_specific_and_supp <- inner_join(supp_table, cluster_3_specific, by = c('SYMBOL' = 'gene')) 
cluster_13_specific_and_supp <- inner_join(supp_table, cluster_13_specific, by = c('SYMBOL' = 'gene'))
cluster_11_specific_and_supp <- inner_join(supp_table, cluster_11_specific, by = c('SYMBOL' = 'gene'))
cluster_22_specific_and_supp <- inner_join(supp_table, cluster_22_specific, by = c('SYMBOL' = 'gene'))
cluster_27_specific_and_supp <- inner_join(supp_table, cluster_27_specific, by = c('SYMBOL' = 'gene'))

#write.csv(cluster_3_specific_and_supp, '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/Comparing_clust_3_11_13_22_27/cluster_3_specific_and_supp.csv', row.names = F)
#write.csv(cluster_11_specific_and_supp, '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/Comparing_clust_3_11_13_22_27/cluster_11_specific_and_supp.csv', row.names = F)
#write.csv(cluster_13_specific_and_supp, '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/Comparing_clust_3_11_13_22_27/cluster_13_specific_and_supp.csv', row.names = F)
#write.csv(cluster_22_specific_and_supp, '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/Comparing_clust_3_11_13_22_27/cluster_22_specific_and_supp.csv', row.names = F)
#write.csv(cluster_27_specific_and_supp, '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/Comparing_clust_3_11_13_22_27/cluster_27_specific_and_supp.csv', row.names = F)

#write.csv(cluster_3_specific, '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/Comparing_clust_3_11_13_22_27/cluster_3_specific.csv', row.names = F)
#write.csv(cluster_11_specific, '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/Comparing_clust_3_11_13_22_27/cluster_11_specific.csv', row.names = F)
#write.csv(cluster_13_specific, '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/Comparing_clust_3_11_13_22_27/cluster_13_specific.csv', row.names = F)
#write.csv(cluster_22_specific, '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/Comparing_clust_3_11_13_22_27/cluster_22_specific.csv', row.names = F)
#write.csv(cluster_27_specific, '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/Comparing_clust_3_11_13_22_27/cluster_27_specific.csv', row.names = F)

```

```{r}
#Let's also look at all the markers in clusters 3, 11, 13, 22, and 27 -- is there any KEGG enrichment there? 
library(clusterProfiler)
install.packages('ggnewscale')

#get the Ensembl mart:
mouse_biomart <- useMart("ensembl", 
                         dataset = "mmusculus_gene_ensembl")

entrez_universe <- getBM(mart = mouse_biomart, 
                                attributes = c("mgi_description", "mgi_symbol", "entrezgene_id"), 
                                values = rownames(all_data.combined), 
                                filters = "mgi_symbol")

entrez_universe$entrezgene_id < as.character(entrez_universe$entrezgene_id)
```

```{r}

mouse_biomart_entrez_3 <- getBM(mart = mouse_biomart, 
                                attributes = c("mgi_description", "mgi_symbol", "entrezgene_id"), 
                                values = cluster_3_specific$gene, 
                                filters = "mgi_symbol")
gseKEGG_3 <- enrichKEGG(gene = mouse_biomart_entrez_3$entrezgene_id, 
                        organism = 'mmu',
                        universe = as.character(entrez_universe$entrezgene_id))


mouse_biomart_entrez_11 <- getBM(mart = mouse_biomart, 
                                attributes = c("mgi_description", "mgi_symbol", "entrezgene_id"), 
                                values = cluster_11_specific$gene, 
                                filters = "mgi_symbol")
gseKEGG_11 <- enrichKEGG(gene = mouse_biomart_entrez_11$entrezgene_id, 
                        organism = 'mmu',
                        universe = as.character(entrez_universe$entrezgene_id))


mouse_biomart_entrez_13 <- getBM(mart = mouse_biomart, 
                                attributes = c("mgi_description", "mgi_symbol", "entrezgene_id"), 
                                values = cluster_13_specific$gene, 
                                filters = "mgi_symbol")
gseKEGG_13 <- enrichKEGG(gene = mouse_biomart_entrez_13$entrezgene_id, 
                        organism = 'mmu',
                        universe = as.character(entrez_universe$entrezgene_id))


mouse_biomart_entrez_22 <- getBM(mart = mouse_biomart, 
                                attributes = c("mgi_description", "mgi_symbol", "entrezgene_id"), 
                                values = cluster_22_specific$gene, 
                                filters = "mgi_symbol")
gseKEGG_22 <- enrichKEGG(gene = mouse_biomart_entrez_22$entrezgene_id, 
                        organism = 'mmu',
                        universe = as.character(entrez_universe$entrezgene_id))


mouse_biomart_entrez_27 <- getBM(mart = mouse_biomart, 
                                attributes = c("mgi_description", "mgi_symbol", "entrezgene_id"), 
                                values = cluster_27_specific$gene, 
                                filters = "mgi_symbol")
gseKEGG_27 <- enrichKEGG(gene = mouse_biomart_entrez_27$entrezgene_id, 
                        organism = 'mmu',
                        universe = as.character(entrez_universe$entrezgene_id))
```
```{r}
#BiocManager::install("org.Mm.eg.db")

gseKEGG_13 <- setReadable(gseKEGG_13, 'org.Mm.eg.db', 'ENTREZID')
gseKEGG_13_dotplot <- dotplot(gseKEGG_13)
gseKEGG_13_cnetplot <- cnetplot(gseKEGG_13)
gseKEGG_13_plot <- cowplot::plot_grid(gseKEGG_13_dotplot, gseKEGG_13_cnetplot)
gseKEGG_13_plot$data$geneID
gseKEGG_13_plot$data$Description
ggsave(plot = gseKEGG_13_plot, '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/cluster13_kegg.png', height=4, width=10)
write.csv(data.frame(gseKEGG_13), '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/cluster13_kegg.csv')

```

```{r}
gseKEGG_22 <- setReadable(gseKEGG_22, 'org.Mm.eg.db', 'ENTREZID')
gseKEGG_22_dotplot <- dotplot(gseKEGG_22, showCategory = 7)
gseKEGG_22_cnetplot <- cnetplot(gseKEGG_22)
gseKEGG_22_plot <- cowplot::plot_grid(gseKEGG_22_dotplot, gseKEGG_22_cnetplot, ncol = 1)
gseKEGG_22_plot$data$geneID
gseKEGG_22_plot$data$Description
ggsave(plot = gseKEGG_22_plot, '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/cluster22_kegg.png', height=15, width=15)
write.csv(data.frame(gseKEGG_22), '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/cluster22_kegg.csv')
```

```{r}
gseKEGG_27 <- setReadable(gseKEGG_27, 'org.Mm.eg.db', 'ENTREZID')
gseKEGG_27_dotplot <- dotplot(gseKEGG_27)
gseKEGG_27_cnetplot <- cnetplot(gseKEGG_27)
gseKEGG_27_plot <- cowplot::plot_grid(gseKEGG_27_dotplot, gseKEGG_27_cnetplot)
gseKEGG_27_plot$data$geneID
gseKEGG_27_plot$data$Description
ggsave(plot = gseKEGG_27_plot, '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/cluster27_kegg.png', height=4, width=15)
write.csv(data.frame(gseKEGG_27), '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/cluster27_kegg.csv')
```
BiocManager::install("pathview")
library('pathview')
```{r}
covid_genes <- data.frame(gseKEGG_22) %>% dplyr::filter(ID == 'mmu05171') 
covid_genes <- covid_genes$geneID
covid_genes <- unlist(strsplit(covid_genes, '\\/'))
covid_genes <- data.frame(covid_genes)
colnames(covid_genes) <- 'mgi_symbol'
covid_ids <- left_join(covid_genes, entrez_universe)
setwd('/Users/jwalla12/Repositories/dennery_scrna/scRNAseq')
pathview(gene.data = covid_ids$mgi_symbol,
         pathway.id = "mmu05171",
         species = "mmu",
         gene.idtype = "SYMBOL",
         gene.annotpkg = "org.Mm.eg.db"
         )

```







```{r}
#Idents(all_data.combined) <- all_data.combined$seurat_clusters
#expression_mat <- AverageExpression(all_data.combined, assays = 'SCT', add.ident = 'orig.ident')
```



```{r Make some specific comparisons...}

#Idents(all_data.combined) <- all_data.combined$orig.ident
#all_data.combined.AirD60_vs_O2D60 <- FindMarkers(all_data.combined, assay = 'SCT', ident.1 = 'AirD60', ident.2 = 'O2D60')





```


```{r Look at specific pathways of interest...}
#filter to get the genes in each of these kegg pathways or descrption keywords
glycolysis <- all_genes_bm %>% 
  filter(str_detect(kegg_enzyme, "^00010+"))

tca <- all_genes_bm %>% 
  filter(str_detect(kegg_enzyme, "^00020+"))

hmox <- all_genes_bm %>%
  filter(str_detect(mgi_description, "heme oxygenase"))

#make a dotplot of the hmox genes for each cluster:
hmox_dotplots <- lapply(seq_along(seurat.list), function(i) DotPlot(seurat.list[[i]], assay = 'SCT', features = c(hmox$mgi_symbol)) + RotatedAxis() + ggtitle(paste0("Heme oxygenase genes - Seurat Cluster ", names(seurat.list[i]))))

glycolysis_dotplots <- lapply(seq_along(seurat.list), function(i) DotPlot(seurat.list[[i]], assay = 'SCT', features = c(glycolysis$mgi_symbol)) + RotatedAxis() + ggtitle(paste0("Glycolysis genes - Seurat Cluster ", names(seurat.list[i]))))

tca_dotplots <- lapply(seq_along(seurat.list), function(i) DotPlot(seurat.list[[i]], assay = 'SCT', features = c(unique(tca$mgi_symbol))) + RotatedAxis() + ggtitle(paste0("TCA genes - Seurat Cluster ", names(seurat.list[i]))))

hmox_dotplots
glycolysis_dotplots
tca_dotplots
#save the dotplots
#lapply(seq_along(hmox_dotplots), function(i) ggsave(plot = hmox_dotplots[[i]], file = paste0("/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/figures/hmox_dotplots/cluster_", i-1, "_hmox_dotplot.pdf"), width = 11,  height = 7))

#save the dotplots
#lapply(seq_along(glycolysis_dotplots), function(i) ggsave(plot = glycolysis_dotplots[[i]], file = paste0("/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/figures/glycolysis_dotplots/cluster_", i-1, "_glycolysis_dotplot.pdf"), width = 15,  height = 7))

#save the dotplots
#lapply(seq_along(tca_dotplots), function(i) ggsave(plot = tca_dotplots[[i]], file = paste0("/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/figures/tca_dotplots/cluster_", i-1, "_tca_dotplot.pdf"), width = 11,  height = 7))
```









```{r}
#AT1_allmarkers <- FindAllMarkers(all_data.combined_AT1)
#write.table(AT1_allmarkers, '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/AT1_allmarkers.txt')

counts <- data.frame(all_data.combined@assays$SCT@counts) #these are the SCTransform corrected counts
counts$gene <- rownames(counts)
counts_melt <- melt(counts, id.vars = 'gene')
counts_melt <- tidyr::separate(data= counts_melt, col = variable, into = c('orig.ident', 'umi'), sep = "_", remove = F) %>% tidyr::separate(col = orig.ident, into = c('treat', 'time'), sep = "D", remove = F)

counts_melt$treat <- as.factor(counts_melt$treat) %>% relevel(ref = 'Air')
counts_melt$time <- as.factor(counts_melt$time) %>% relevel(ref = '7')

counts_melt <- rename(counts_melt, 'value' = 'counts')

# Build the linear model
model  <- lm(counts ~ treat*time,
             data = data.frame(counts_melt))

# the residuals are not normally distributed (p-value is significant)
counts_melt %>%
  group_by(treat, time) %>%
  slice_sample(n = 5000) %>%
  shapiro_test(counts)

counts_melt %>%
  slice_sample(n = 5000) %>%
  levene_test(counts ~ treat*time)

counts_melt_aov <- counts_melt %>%
  slice_sample(n = 5000) %>%
  anova_test(counts ~ treat*time)

counts_melt_aov

counts_melt %>%
  group_by(treat)%>%
  anova_test(counts ~ time, error = model)
```


```{r}
#AT1_allmarkers <- FindAllMarkers(all_data.combined_AT1)
#write.table(AT1_allmarkers, '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/AT1_allmarkers.txt')

AT1_counts <- data.frame(all_data.combined_AT1@assays$SCT@counts) #these are the SCTransform corrected counts
AT1_counts$gene <- rownames(AT1_counts)
AT1_counts_melt <- melt(AT1_counts, id.vars = 'gene')
AT1_counts_melt <- tidyr::separate(data= AT1_counts_melt, col = variable, into = c('orig.ident', 'umi'), sep = "_", remove = F) %>% tidyr::separate(col = orig.ident, into = c('treat', 'time'), sep = "D", remove = F)

AT1_counts_melt$treat <- as.factor(AT1_counts_melt$treat) %>% relevel(ref = 'Air')
AT1_counts_melt$time <- as.factor(AT1_counts_melt$time) %>% relevel(ref = '7')

AT1_counts_melt <- rename(AT1_counts_melt, 'value' = 'counts')

# Build the linear model
AT1_model  <- lm(counts ~ treat*time,
             data = data.frame(AT1_counts_melt))

# the residuals are not normally distributed (p-value is significant)
AT1_counts_melt %>%
  group_by(treat, time) %>%
  slice_sample(n = 5000) %>%
  shapiro_test(counts)

AT1_counts_melt %>%
  slice_sample(n = 5000) %>%
  levene_test(counts ~ treat*time)

AT1_counts_melt_aov <- AT1_counts_melt %>%
  slice_sample(n = 5000) %>%
  anova_test(counts ~ treat*time)

AT1_counts_melt_aov

AT1_counts_melt %>%
  group_by(treat)%>%
  anova_test(counts ~ time, error = AT1_model)
```
# So, there are definitely some interactions between time and treatment.


#DefaultAssay(all_data.combined_AT1) <- 'SCT'
```{r}
#O2D7_v_AirD7 <- FindMarkers(all_data.combined, ident.1 = "O2D7", ident.2 = "AirD7", group.by = 'orig.ident', assay = 'SCT')
```


```{r}
#O2D7_v_AirD7_AT1 <- FindMarkers(all_data.combined, ident.1 = "O2D7-25", ident.2 = "AirD7-25", group.by = 'ident_clust')
```

```{r}
#AllMarkers <- FindAllMarkers(all_data.combined, group.by = 'orig.ident', test.use = 'bimod')

```

```{r}
#save.image(file = "/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/dennery.RData")
```

```{r}
#load(file = "/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/dennery.RData")
```

```{r}
#library(monocle)

```

```{r}
#DefaultAssay(all_data.combined) <- 'RNA'

```

```{r}

##all_data.combined_gene_metadata <- data.frame(rownames(all_data.combined@assays$RNA))
#colnames(all_data.combined_gene_metadata) <- 'gene_short_name'
#row.names(all_data.combined_gene_metadata) <- all_data.combined_gene_metadata$gene_short_name
```


```{r}
#ggsave(file = '/Users/jwalla12/Repositories/dennery_scrna/scRNAseq/figures_and_tables/figures/qc_figures/supplemental_figure_1_filtering_metrics.png', plot = wrapped_filtering)
```


#We identified a total of  45 clusters of lung cells along with corresponding marker genes (Fig 2Ax, Table 1x). 

#This included 19 cell types  that included epithelial cells, endothelial cells, mesenchymal cells, and immune cells (Fig 2B and 2Cx, Table 2x)

